---
title: "Explore K-12 Missing Trip Ends"
author: "David Ory"
output:
  html_document:
    theme: cosmo
    toc: yes
runtime: shiny
---

## Administration

#### _TODO_
1.  Join with enrollment data to check for mismatches

#### Purpose
When estimating size terms, we compare trip ends to likely trip attractors.  For school trips, we compare trip ends for day care, pre-school, grade, middle, and high school from the California Household Travel Survey with K-12 employment estimates.  When doing these comparisons at the TAZ geography, many trip ends are attracted to TAZs that have no K-12 employment.  The purpose of this exploration is to look at these 'un-matched' trip ends and determine what has gone wrong.

## Overhead

#### Libraries
```{r overhead}
library(knitr)
library(reshape2)
suppressMessages(library(dplyr))
library(stringr)
```

```{r config, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)

```

## Procedure

#### Read in the data
```{r data-read}

data.nomatch <- read.csv('D:/files/My Box Files/Travel Model Two Development/Demand/CHTS Processing/_WORKING/School Trip Ends Exploration/schoolTrips_NoK12Employment.csv', header = TRUE)

```

## Interactives

### High School

```{r shiny_table_highschool, echo=FALSE, cache=FALSE}

textInput("NumberHighSchool", "Number of high school rows to show:", "15")

renderTable({
  
  data.hs.sum <- data.nomatch %.%
  filter(hs_trip == 1) %.%
  group_by(PNAME, taz) %.%
  summarise(Count = n()) %.%
  arrange(-Count)

  head(data.hs.sum, n = input$NumberHighSchool)
   
  })

```

### K-8 School

```{r shiny_table_k8school, echo=FALSE, cache=FALSE}

textInput("NumberK8School", "Number of grade or middle school rows to show:", "15")

renderTable({
  
  data.k8.sum <- data.nomatch %.%
  filter(k8_trip == 1) %.%
  group_by(PNAME, taz) %.%
  summarise(Count = n()) %.%
  arrange(-Count)

  head(data.k8.sum, n = input$NumberK8School)
   
  })

```


## Log Manual Checks

#### High Schools
Top Five TAZ Location Checks
1.  Homestead HS, 200297, Check
2.  Westmont HS, 200358, Check
3.  Burlingame HS, 100134, Check
4.  Los Altos HS, 200974, Check
5.  Adrian Wilcox HS, 200380, Check

#### K8 Schools
Top Five TAZ Location Checks
1.  Nelda Mundy, 500184, Check
2.  Piedmont Middle School, 300954, Check
3.  Borel Middle School, 100330, Check
4.  Green Elementary School, 300245 (300784), Fail
5.  Apple Blossom School, 700033, Check

