---
title: "Mandatory Tour Frequency"
output:
  html_document:
    df_print: paged
---

# Overhead
```{r overhead, include = FALSE}
packages_vector <- c("tidyverse",
                     "kableExtra")

need_to_install <- packages_vector[!(packages_vector %in% installed.packages()[,"Package"])]

if (length(need_to_install)) install.packages(need_to_install)

for (package in packages_vector) {
  library(package, character.only = TRUE)
}

```

# Remote I/O
```{r remote-io}
interim_dir <- "../output/"

person_filename <- paste0(interim_dir, "final_persons.csv")
tour_filename <- paste0(interim_dir, "final_tours.csv")

```

# Parameters
```{r parameters}
ptype_dict <- tibble(ptype = seq(1,8),
                     label = c("Full-time Worker",
                               "Part-time Worker",
                               "University Student",
                               "Unemployed",
                               "Retired",
                               "Driving age Student",
                               "Non-driving age Student",
                               "Preschool Student"))
```


# Data Reads
Minor note: bloated output -- see `parking-subsidy.Rmd` note. 
```{r read}
person_df <- read_csv(person_filename, col_types = cols(
  person_id = col_double(),
  household_id = col_double(),
  age = col_double(),
  PNUM = col_double(),
  sex = col_double(),
  pemploy = col_double(),
  pstudent = col_double(),
  ptype = col_double(),
  educ = col_double(),
  naics2_original_code = col_character(),
  soc2 = col_double(),
  age_16_to_19 = col_logical(),
  age_16_p = col_logical(),
  adult = col_logical(),
  male = col_logical(),
  female = col_logical(),
  has_non_worker = col_logical(),
  has_retiree = col_logical(),
  has_preschool_kid = col_logical(),
  has_driving_kid = col_logical(),
  has_school_kid = col_logical(),
  has_full_time = col_logical(),
  has_part_time = col_logical(),
  has_university = col_logical(),
  student_is_employed = col_logical(),
  nonstudent_to_school = col_logical(),
  is_student = col_logical(),
  is_preschool = col_logical(),
  is_gradeschool = col_logical(),
  is_highschool = col_logical(),
  is_university = col_logical(),
  school_segment = col_double(),
  is_worker = col_logical(),
  is_fulltime_worker = col_logical(),
  is_parttime_worker = col_logical(),
  is_internal_worker = col_logical(),
  is_external_worker = col_logical(),
  home_zone_id = col_double(),
  time_factor_work = col_double(),
  time_factor_nonwork = col_double(),
  naics_code = col_double(),
  occupation = col_character(),
  is_income_less25K = col_logical(),
  is_income_25K_to_60K = col_logical(),
  is_income_60K_to_120K = col_logical(),
  is_income_greater60K = col_logical(),
  is_income_greater120K = col_logical(),
  is_non_worker_in_HH = col_logical(),
  is_all_adults_full_time_workers = col_logical(),
  is_pre_drive_child_in_HH = col_logical(),
  work_from_home = col_logical(),
  is_out_of_home_worker = col_logical(),
  external_workplace_zone_id = col_double(),
  external_workplace_location_logsum = col_double(),
  external_workplace_modechoice_logsum = col_double(),
  school_zone_id = col_double(),
  school_location_logsum = col_double(),
  school_modechoice_logsum = col_double(),
  distance_to_school = col_double(),
  roundtrip_auto_time_to_school = col_double(),
  workplace_zone_id = col_double(),
  workplace_location_logsum = col_double(),
  workplace_modechoice_logsum = col_double(),
  distance_to_work = col_double(),
  workplace_in_cbd = col_logical(),
  work_zone_area_type = col_double(),
  auto_time_home_to_work = col_double(),
  roundtrip_auto_time_to_work = col_double(),
  work_auto_savings = col_double(),
  exp_daily_work = col_double(),
  non_toll_time_work = col_double(),
  toll_time_work = col_double(),
  toll_dist_work = col_double(),
  toll_cost_work = col_double(),
  toll_travel_time_savings_work = col_double(),
  transit_pass_subsidy = col_double(),
  transit_pass_ownership = col_double(),
  free_parking_at_work = col_logical(),
  telecommute_frequency = col_character(),
  cdap_activity = col_character(),
  travel_active = col_logical(),
  under16_not_at_school = col_logical(),
  has_preschool_kid_at_home = col_logical(),
  has_school_kid_at_home = col_logical(),
  mandatory_tour_frequency = col_character(),
  work_and_school_and_worker = col_logical(),
  work_and_school_and_student = col_logical(),
  num_mand = col_double(),
  num_work_tours = col_double(),
  has_pre_school_child_with_mandatory = col_logical(),
  has_driving_age_child_with_mandatory = col_logical(),
  num_joint_tours = col_double(),
  non_mandatory_tour_frequency = col_double(),
  num_non_mand = col_double(),
  num_escort_tours = col_double(),
  num_eatout_tours = col_double(),
  num_shop_tours = col_double(),
  num_maint_tours = col_double(),
  num_discr_tours = col_double(),
  num_social_tours = col_double(),
  num_non_escort_tours = col_double(),
  num_shop_maint_tours = col_double(),
  num_shop_maint_escort_tours = col_double(),
  num_add_shop_maint_tours = col_double(),
  num_soc_discr_tours = col_double(),
  num_add_soc_discr_tours = col_double(),
  model = col_character()
))

tour_df <- read_csv(tour_filename, col_types = cols(
  tour_id = col_double(),
  person_id = col_double(),
  tour_type = col_character(),
  tour_type_count = col_double(),
  tour_type_num = col_double(),
  tour_num = col_double(),
  tour_count = col_double(),
  tour_category = col_character(),
  number_of_participants = col_double(),
  destination = col_double(),
  origin = col_double(),
  household_id = col_double(),
  start = col_double(),
  end = col_double(),
  duration = col_double(),
  school_esc_outbound = col_character(),
  school_esc_inbound = col_character(),
  num_escortees = col_double(),
  tdd = col_double(),
  tour_id_temp = col_double(),
  composition = col_character(),
  is_external_tour = col_logical(),
  is_internal_tour = col_logical(),
  destination_logsum = col_double(),
  vehicle_occup_1 = col_character(),
  vehicle_occup_2 = col_character(),
  vehicle_occup_3.5 = col_character(),
  tour_mode = col_character(),
  mode_choice_logsum = col_double(),
  selected_vehicle = col_character(),
  atwork_subtour_frequency = col_character(),
  parent_tour_id = col_double(),
  stop_frequency = col_character(),
  primary_purpose = col_character(),
  model = col_character()
))
```

# Reductions 
```{r reductions}
temp_df <- tour_df %>%
  select(tour_id, person_id, tour_category, tour_type) %>%
  filter(tour_category == "mandatory") %>%
  group_by(person_id, tour_type) %>%
  summarise(mandatory_tours = n(), .groups = "drop") %>%
  mutate(choice = paste0(tour_type, "_", mandatory_tours))

working_df <- temp_df %>%
  group_by(person_id) %>%
  summarise(count = n(), .groups = "drop") %>%
  filter(count > 1) %>%
  mutate(choice_update = "work_and_school") %>%
  select(person_id, choice_update)

out_df <- left_join(temp_df, working_df, by = c("person_id")) %>%
  mutate(choice = if_else(is.na(choice_update), choice, choice_update)) %>%
  distinct(person_id, choice) %>%
  left_join(select(person_df, person_id, ptype, cdap_activity), ., by = c("person_id")) %>%
  mutate(choice = if_else(is.na(choice), "none", choice)) %>%
  left_join(., ptype_dict, by = c("ptype")) %>%
  rename(person_type = label)

summary_df <- out_df %>%
  group_by(ptype, person_type, choice) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(ptype, person_type) %>%
  mutate(share = count/sum(count)) %>%
  ungroup() %>%
  arrange(ptype) %>%
  select(person_type, choice, share) %>%
  pivot_wider(names_from = choice, values_from = share, values_fill = 0.0)

summary_df %>%
  kbl() %>%
  kable_styling()

summary_cdap_df <- person_df %>%
  left_join(., ptype_dict, by = c("ptype")) %>%
  rename(person_type = label) %>%
  group_by(ptype, person_type, cdap_activity) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(ptype, person_type) %>%
  mutate(share = count/sum(count)) %>%
  ungroup() %>%
  arrange(ptype) %>%
  select(person_type, cdap_activity, share)

summary_cdap_df %>%
  kbl() %>%
  kable_styling()


```
The model does show the right types of categories for each of the person types. It seems to have a much higher than expected share of non-mandatory and home activity patterns for workers and students, which suggests we have a version prior to additional calibration? But the implementation appears correct. ActivitySim files reviewed include:

- `mandatory_tour_frequency_alternatives.csv`
- `mandatory_tour_frequency_coeffs.csv`
- `mandatory_tour_frequency.csv`
- `mandatory_tour_frequency.yaml`
- `annotate_persons_mtf.csv`



###

