---
title: "Parking Subsidy"
output:
  html_document:
    df_print: paged
---

# Overhead
```{r overhead, include = FALSE}
packages_vector <- c("tidyverse",
                     "kableExtra")

need_to_install <- packages_vector[!(packages_vector %in% installed.packages()[,"Package"])]

if (length(need_to_install)) install.packages(need_to_install)

for (package in packages_vector) {
  library(package, character.only = TRUE)
}

```

# Remote I/O
```{r remote-io}
interim_dir <- "../output/"

person_filename <- paste0(interim_dir, "final_persons.csv")
trips_filename <- paste0(interim_dir, "final_trips.csv")

temp_trips_filename <- paste0(interim_dir, "final_trips.RDS")

```

# Parameters
```{r parameters}
ptype_dict <- tibble(ptype = seq(1,8),
                     label = c("Full-time Worker",
                               "Part-time Worker",
                               "University Student",
                               "Unemployed",
                               "Retired",
                               "Driving age Student",
                               "Non-driving age Student",
                               "Preschool Student"))
```


# Data Reads
Minor note: the ActivitySim outputs get bloated with temporary variables that are poorly documented. Using a data model to clean-up and document the output is low-hanging fruit. 
```{r read-persons}
person_df <- read_csv(person_filename, col_types = cols(
  person_id = col_double(),
  household_id = col_double(),
  age = col_double(),
  PNUM = col_double(),
  sex = col_double(),
  pemploy = col_double(),
  pstudent = col_double(),
  ptype = col_double(),
  educ = col_double(),
  naics2_original_code = col_character(),
  soc2 = col_double(),
  age_16_to_19 = col_logical(),
  age_16_p = col_logical(),
  adult = col_logical(),
  male = col_logical(),
  female = col_logical(),
  has_non_worker = col_logical(),
  has_retiree = col_logical(),
  has_preschool_kid = col_logical(),
  has_driving_kid = col_logical(),
  has_school_kid = col_logical(),
  has_full_time = col_logical(),
  has_part_time = col_logical(),
  has_university = col_logical(),
  student_is_employed = col_logical(),
  nonstudent_to_school = col_logical(),
  is_student = col_logical(),
  is_preschool = col_logical(),
  is_gradeschool = col_logical(),
  is_highschool = col_logical(),
  is_university = col_logical(),
  school_segment = col_double(),
  is_worker = col_logical(),
  is_fulltime_worker = col_logical(),
  is_parttime_worker = col_logical(),
  is_internal_worker = col_logical(),
  is_external_worker = col_logical(),
  home_zone_id = col_double(),
  time_factor_work = col_double(),
  time_factor_nonwork = col_double(),
  naics_code = col_double(),
  occupation = col_character(),
  is_income_less25K = col_logical(),
  is_income_25K_to_60K = col_logical(),
  is_income_60K_to_120K = col_logical(),
  is_income_greater60K = col_logical(),
  is_income_greater120K = col_logical(),
  is_non_worker_in_HH = col_logical(),
  is_all_adults_full_time_workers = col_logical(),
  is_pre_drive_child_in_HH = col_logical(),
  work_from_home = col_logical(),
  is_out_of_home_worker = col_logical(),
  external_workplace_zone_id = col_double(),
  external_workplace_location_logsum = col_double(),
  external_workplace_modechoice_logsum = col_double(),
  school_zone_id = col_double(),
  school_location_logsum = col_double(),
  school_modechoice_logsum = col_double(),
  distance_to_school = col_double(),
  roundtrip_auto_time_to_school = col_double(),
  workplace_zone_id = col_double(),
  workplace_location_logsum = col_double(),
  workplace_modechoice_logsum = col_double(),
  distance_to_work = col_double(),
  workplace_in_cbd = col_logical(),
  work_zone_area_type = col_double(),
  auto_time_home_to_work = col_double(),
  roundtrip_auto_time_to_work = col_double(),
  work_auto_savings = col_double(),
  exp_daily_work = col_double(),
  non_toll_time_work = col_double(),
  toll_time_work = col_double(),
  toll_dist_work = col_double(),
  toll_cost_work = col_double(),
  toll_travel_time_savings_work = col_double(),
  transit_pass_subsidy = col_double(),
  transit_pass_ownership = col_double(),
  free_parking_at_work = col_logical(),
  telecommute_frequency = col_character(),
  cdap_activity = col_character(),
  travel_active = col_logical(),
  under16_not_at_school = col_logical(),
  has_preschool_kid_at_home = col_logical(),
  has_school_kid_at_home = col_logical(),
  mandatory_tour_frequency = col_character(),
  work_and_school_and_worker = col_logical(),
  work_and_school_and_student = col_logical(),
  num_mand = col_double(),
  num_work_tours = col_double(),
  has_pre_school_child_with_mandatory = col_logical(),
  has_driving_age_child_with_mandatory = col_logical(),
  num_joint_tours = col_double(),
  non_mandatory_tour_frequency = col_double(),
  num_non_mand = col_double(),
  num_escort_tours = col_double(),
  num_eatout_tours = col_double(),
  num_shop_tours = col_double(),
  num_maint_tours = col_double(),
  num_discr_tours = col_double(),
  num_social_tours = col_double(),
  num_non_escort_tours = col_double(),
  num_shop_maint_tours = col_double(),
  num_shop_maint_escort_tours = col_double(),
  num_add_shop_maint_tours = col_double(),
  num_soc_discr_tours = col_double(),
  num_add_soc_discr_tours = col_double(),
  model = col_character()
))

```

```{r read-trips, eval=FALSE}
trips_df <- read_csv(trips_filename, col_types = cols(
  trip_id = col_double(),
  person_id = col_double(),
  household_id = col_double(),
  primary_purpose = col_character(),
  trip_num = col_double(),
  outbound = col_logical(),
  trip_count = col_double(),
  destination = col_double(),
  origin = col_double(),
  tour_id = col_double(),
  escort_participants = col_character(),
  school_escort_direction = col_character(),
  purpose = col_character(),
  destination_logsum = col_double(),
  depart = col_double(),
  trip_mode = col_character(),
  mode_choice_logsum = col_double(),
  vot_da = col_double(),
  vot_s2 = col_double(),
  vot_s3 = col_double(),
  ownsTransponder = col_logical(),
  parking_zone = col_double(),
  trip_period = col_character(),
  tour_participants = col_double(),
  distance = col_double(),
  is_ea = col_logical(),
  is_am = col_logical(),
  is_md = col_logical(),
  is_pm = col_logical(),
  is_ev = col_logical(),
  ownTrp = col_logical(),
  vot1 = col_double(),
  vot2 = col_double(),
  vot3 = col_double(),
  inbound = col_logical(),
  DRIVEALONE_NOTR_EA_LOW = col_double(),
  SHARED2_NOTR_EA_LOW = col_double(),
  SHARED3_NOTR_EA_LOW = col_double(),
  DRIVEALONE_TR_EA_LOW = col_double(),
  SHARED2_TR_EA_LOW = col_double(),
  SHARED3_TR_EA_LOW = col_double(),
  DRIVEALONE_NOTR_EA_MED = col_double(),
  SHARED2_NOTR_EA_MED = col_double(),
  SHARED3_NOTR_EA_MED = col_double(),
  DRIVEALONE_TR_EA_MED = col_double(),
  SHARED2_TR_EA_MED = col_double(),
  SHARED3_TR_EA_MED = col_double(),
  DRIVEALONE_NOTR_EA_HIGH = col_double(),
  SHARED2_NOTR_EA_HIGH = col_double(),
  SHARED3_NOTR_EA_HIGH = col_double(),
  DRIVEALONE_TR_EA_HIGH = col_double(),
  SHARED2_TR_EA_HIGH = col_double(),
  SHARED3_TR_EA_HIGH = col_double(),
  WALK_LOC_EA = col_double(),
  WALK_PRM_EA = col_double(),
  WALK_MIX_EA = col_double(),
  PNR_LOCOUT_EA = col_double(),
  PNR_PRMOUT_EA = col_double(),
  PNR_MIXOUT_EA = col_double(),
  KNR_LOCOUT_EA = col_double(),
  KNR_PRMOUT_EA = col_double(),
  KNR_MIXOUT_EA = col_double(),
  TNC_LOCOUT_EA = col_double(),
  TNC_PRMOUT_EA = col_double(),
  TNC_MIXOUT_EA = col_double(),
  PNR_LOCIN_EA = col_double(),
  PNR_PRMIN_EA = col_double(),
  PNR_MIXIN_EA = col_double(),
  KNR_LOCIN_EA = col_double(),
  KNR_PRMIN_EA = col_double(),
  KNR_MIXIN_EA = col_double(),
  TNC_LOCIN_EA = col_double(),
  TNC_PRMIN_EA = col_double(),
  TNC_MIXIN_EA = col_double(),
  BIKE_EA = col_double(),
  WALK_EA = col_double(),
  DRIVEALONE_NOTR_AM_LOW = col_double(),
  SHARED2_NOTR_AM_LOW = col_double(),
  SHARED3_NOTR_AM_LOW = col_double(),
  DRIVEALONE_TR_AM_LOW = col_double(),
  SHARED2_TR_AM_LOW = col_double(),
  SHARED3_TR_AM_LOW = col_double(),
  DRIVEALONE_NOTR_AM_MED = col_double(),
  SHARED2_NOTR_AM_MED = col_double(),
  SHARED3_NOTR_AM_MED = col_double(),
  DRIVEALONE_TR_AM_MED = col_double(),
  SHARED2_TR_AM_MED = col_double(),
  SHARED3_TR_AM_MED = col_double(),
  DRIVEALONE_NOTR_AM_HIGH = col_double(),
  SHARED2_NOTR_AM_HIGH = col_double(),
  SHARED3_NOTR_AM_HIGH = col_double(),
  DRIVEALONE_TR_AM_HIGH = col_double(),
  SHARED2_TR_AM_HIGH = col_double(),
  SHARED3_TR_AM_HIGH = col_double(),
  WALK_LOC_AM = col_double(),
  WALK_PRM_AM = col_double(),
  WALK_MIX_AM = col_double(),
  PNR_LOCOUT_AM = col_double(),
  PNR_PRMOUT_AM = col_double(),
  PNR_MIXOUT_AM = col_double(),
  KNR_LOCOUT_AM = col_double(),
  KNR_PRMOUT_AM = col_double(),
  KNR_MIXOUT_AM = col_double(),
  TNC_LOCOUT_AM = col_double(),
  TNC_PRMOUT_AM = col_double(),
  TNC_MIXOUT_AM = col_double(),
  PNR_LOCIN_AM = col_double(),
  PNR_PRMIN_AM = col_double(),
  PNR_MIXIN_AM = col_double(),
  KNR_LOCIN_AM = col_double(),
  KNR_PRMIN_AM = col_double(),
  KNR_MIXIN_AM = col_double(),
  TNC_LOCIN_AM = col_double(),
  TNC_PRMIN_AM = col_double(),
  TNC_MIXIN_AM = col_double(),
  BIKE_AM = col_double(),
  WALK_AM = col_double(),
  DRIVEALONE_NOTR_MD_LOW = col_double(),
  SHARED2_NOTR_MD_LOW = col_double(),
  SHARED3_NOTR_MD_LOW = col_double(),
  DRIVEALONE_TR_MD_LOW = col_double(),
  SHARED2_TR_MD_LOW = col_double(),
  SHARED3_TR_MD_LOW = col_double(),
  DRIVEALONE_NOTR_MD_MED = col_double(),
  SHARED2_NOTR_MD_MED = col_double(),
  SHARED3_NOTR_MD_MED = col_double(),
  DRIVEALONE_TR_MD_MED = col_double(),
  SHARED2_TR_MD_MED = col_double(),
  SHARED3_TR_MD_MED = col_double(),
  DRIVEALONE_NOTR_MD_HIGH = col_double(),
  SHARED2_NOTR_MD_HIGH = col_double(),
  SHARED3_NOTR_MD_HIGH = col_double(),
  DRIVEALONE_TR_MD_HIGH = col_double(),
  SHARED2_TR_MD_HIGH = col_double(),
  SHARED3_TR_MD_HIGH = col_double(),
  WALK_LOC_MD = col_double(),
  WALK_PRM_MD = col_double(),
  WALK_MIX_MD = col_double(),
  PNR_LOCOUT_MD = col_double(),
  PNR_PRMOUT_MD = col_double(),
  PNR_MIXOUT_MD = col_double(),
  KNR_LOCOUT_MD = col_double(),
  KNR_PRMOUT_MD = col_double(),
  KNR_MIXOUT_MD = col_double(),
  TNC_LOCOUT_MD = col_double(),
  TNC_PRMOUT_MD = col_double(),
  TNC_MIXOUT_MD = col_double(),
  PNR_LOCIN_MD = col_double(),
  PNR_PRMIN_MD = col_double(),
  PNR_MIXIN_MD = col_double(),
  KNR_LOCIN_MD = col_double(),
  KNR_PRMIN_MD = col_double(),
  KNR_MIXIN_MD = col_double(),
  TNC_LOCIN_MD = col_double(),
  TNC_PRMIN_MD = col_double(),
  TNC_MIXIN_MD = col_double(),
  BIKE_MD = col_double(),
  WALK_MD = col_double(),
  DRIVEALONE_NOTR_PM_LOW = col_double(),
  SHARED2_NOTR_PM_LOW = col_double(),
  SHARED3_NOTR_PM_LOW = col_double(),
  DRIVEALONE_TR_PM_LOW = col_double(),
  SHARED2_TR_PM_LOW = col_double(),
  SHARED3_TR_PM_LOW = col_double(),
  DRIVEALONE_NOTR_PM_MED = col_double(),
  SHARED2_NOTR_PM_MED = col_double(),
  SHARED3_NOTR_PM_MED = col_double(),
  DRIVEALONE_TR_PM_MED = col_double(),
  SHARED2_TR_PM_MED = col_double(),
  SHARED3_TR_PM_MED = col_double(),
  DRIVEALONE_NOTR_PM_HIGH = col_double(),
  SHARED2_NOTR_PM_HIGH = col_double(),
  SHARED3_NOTR_PM_HIGH = col_double(),
  DRIVEALONE_TR_PM_HIGH = col_double(),
  SHARED2_TR_PM_HIGH = col_double(),
  SHARED3_TR_PM_HIGH = col_double(),
  WALK_LOC_PM = col_double(),
  WALK_PRM_PM = col_double(),
  WALK_MIX_PM = col_double(),
  PNR_LOCOUT_PM = col_double(),
  PNR_PRMOUT_PM = col_double(),
  PNR_MIXOUT_PM = col_double(),
  KNR_LOCOUT_PM = col_double(),
  KNR_PRMOUT_PM = col_double(),
  KNR_MIXOUT_PM = col_double(),
  TNC_LOCOUT_PM = col_double(),
  TNC_PRMOUT_PM = col_double(),
  TNC_MIXOUT_PM = col_double(),
  PNR_LOCIN_PM = col_double(),
  PNR_PRMIN_PM = col_double(),
  PNR_MIXIN_PM = col_double(),
  KNR_LOCIN_PM = col_double(),
  KNR_PRMIN_PM = col_double(),
  KNR_MIXIN_PM = col_double(),
  TNC_LOCIN_PM = col_double(),
  TNC_PRMIN_PM = col_double(),
  TNC_MIXIN_PM = col_double(),
  BIKE_PM = col_double(),
  WALK_PM = col_double(),
  DRIVEALONE_NOTR_EV_LOW = col_double(),
  SHARED2_NOTR_EV_LOW = col_double(),
  SHARED3_NOTR_EV_LOW = col_double(),
  DRIVEALONE_TR_EV_LOW = col_double(),
  SHARED2_TR_EV_LOW = col_double(),
  SHARED3_TR_EV_LOW = col_double(),
  DRIVEALONE_NOTR_EV_MED = col_double(),
  SHARED2_NOTR_EV_MED = col_double(),
  SHARED3_NOTR_EV_MED = col_double(),
  DRIVEALONE_TR_EV_MED = col_double(),
  SHARED2_TR_EV_MED = col_double(),
  SHARED3_TR_EV_MED = col_double(),
  DRIVEALONE_NOTR_EV_HIGH = col_double(),
  SHARED2_NOTR_EV_HIGH = col_double(),
  SHARED3_NOTR_EV_HIGH = col_double(),
  DRIVEALONE_TR_EV_HIGH = col_double(),
  SHARED2_TR_EV_HIGH = col_double(),
  SHARED3_TR_EV_HIGH = col_double(),
  WALK_LOC_EV = col_double(),
  WALK_PRM_EV = col_double(),
  WALK_MIX_EV = col_double(),
  PNR_LOCOUT_EV = col_double(),
  PNR_PRMOUT_EV = col_double(),
  PNR_MIXOUT_EV = col_double(),
  KNR_LOCOUT_EV = col_double(),
  KNR_PRMOUT_EV = col_double(),
  KNR_MIXOUT_EV = col_double(),
  TNC_LOCOUT_EV = col_double(),
  TNC_PRMOUT_EV = col_double(),
  TNC_MIXOUT_EV = col_double(),
  PNR_LOCIN_EV = col_double(),
  PNR_PRMIN_EV = col_double(),
  PNR_MIXIN_EV = col_double(),
  KNR_LOCIN_EV = col_double(),
  KNR_PRMIN_EV = col_double(),
  KNR_MIXIN_EV = col_double(),
  TNC_LOCIN_EV = col_double(),
  TNC_PRMIN_EV = col_double(),
  TNC_MIXIN_EV = col_double(),
  BIKE_EV = col_double(),
  WALK_EV = col_double(),
  timeDrive = col_double(),
  distanceDrive = col_double(),
  costTollDrive = col_double(),
  timeTransitInVehicle = col_double(),
  timeRapidTransitInVehicle = col_double(),
  timeExpressBusTransitInVehicle = col_double(),
  timeLocalBusTransitInVehicle = col_double(),
  timeLightRailTransitInVehicle = col_double(),
  timeCommuterRailTransitInVehicle = col_double(),
  timeTransitInitialWait = col_double(),
  costFareTransit = col_double(),
  transfersTransit = col_double(),
  timeBike = col_double(),
  timeWalk = col_double(),
  sample_rate = col_double(),
  otaz = col_double(),
  dtaz = col_double(),
  model = col_character()
))

```


# Temp I/O
```{r temp-io}
# saveRDS(trips_df, temp_trips_filename)
trips_df <- readRDS(temp_trips_filename)
```


# Reductions 
```{r reductions}
summary_df <- person_df %>%
  select(person_id, ptype, free_parking_at_work) %>%
  left_join(., ptype_dict, by = c("ptype")) %>%
  rename(person_type = label) %>%
  mutate(free_parking_at_work = if_else(free_parking_at_work, 
                                        "Free Parking at Work", 
                                        "Not Free Parking at Work")) %>%
  group_by(ptype, person_type, free_parking_at_work) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(ptype, person_type) %>%
  mutate(share = count/sum(count)) %>%
  ungroup() %>%
  arrange(ptype) %>%
  select(person_type, free_parking_at_work, share) %>%
  pivot_wider(
    names_from = free_parking_at_work, 
    values_from = share, 
    values_fill = 0.0)
  

summary_df %>%
  kbl() %>%
  kable_styling()

```
The model does show the right types of categories for each of the person types. The implementation appears correct. ActivitySim files reviewed include:

- `free_parking_annotate_persons_preprocessor.csv`
- `free_parking_coefficients.csv`
-- For SANDAG: still has coefficients specific to SF Bay Area
- `free_parking.csv`: not much intelligence in the spec
- `free_parking.yaml`: appears as expected
- `tour_mode_choice_annotate_preprocessor.csv`: removes the parking cost for those with free parking
- `trip_mode_choice_annotate_trips_preprocessor.csv`: same as tour 

```{r Reductions part 2}
# need to check what choices those with free parking make in regards to parking locations
summary_df <- trips_df %>%
  select(trip_id, person_id, purpose, parking_zone, destination) %>%
  filter(purpose == "work") %>% 
  left_join(
    ., 
    select(person_df, person_id, free_parking_at_work), 
    by = c("person_id")) %>%
  mutate(parking_choice_made = (parking_zone > 0)) %>%
  mutate(parking_zone_is_destination = if_else(
    parking_zone == destination, 
    "Parked at Destination", 
    "Parked and Walked to Destination")) %>%
  mutate(parking_zone_is_destination = if_else(
    parking_choice_made, 
    parking_zone_is_destination, 
    "No Parking Choice Made")) %>%
  mutate(free_parking_at_work = if_else(
    free_parking_at_work, 
    "Free Parking at Work", 
    "Not Free Parking at Work")) %>%
  group_by(free_parking_at_work, parking_zone_is_destination) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(free_parking_at_work) %>%
  mutate(share = count/sum(count)) %>%
  ungroup() %>%
  select(-count) %>%
  pivot_wider(
    names_from = parking_zone_is_destination, 
    values_from = share, 
    values_fill = 0.0)

summary_df %>%
  kbl() %>%
  kable_styling()
  

```

There is a known problem with ActivitySim in which the `destination` column is overwritten with the parking location. See this [pull request](https://github.com/SANDAG/activitysim/pull/64) for additional information. 



###

