---
title: "Stop Frequency"
output:
  html_document:
    df_print: paged
---

# Overhead
```{r overhead, include = FALSE}
packages_vector <- c("tidyverse",
                     "kableExtra")

need_to_install <- packages_vector[!(packages_vector %in% installed.packages()[,"Package"])]

if (length(need_to_install)) install.packages(need_to_install)

for (package in packages_vector) {
  library(package, character.only = TRUE)
}

```

# Remote I/O
```{r remote-io}
interim_dir <- "../output/"

tour_filename <- paste0(interim_dir, "final_tours.csv")

```

# Parameters
```{r parameters}

```


# Data Reads
```{r read}
tour_df <- read_csv(tour_filename, col_types = cols(
  tour_id = col_double(),
  person_id = col_double(),
  tour_type = col_character(),
  tour_type_count = col_double(),
  tour_type_num = col_double(),
  tour_num = col_double(),
  tour_count = col_double(),
  tour_category = col_character(),
  number_of_participants = col_double(),
  destination = col_double(),
  origin = col_double(),
  household_id = col_double(),
  start = col_double(),
  end = col_double(),
  duration = col_double(),
  school_esc_outbound = col_character(),
  school_esc_inbound = col_character(),
  num_escortees = col_double(),
  tdd = col_double(),
  tour_id_temp = col_double(),
  composition = col_character(),
  is_external_tour = col_logical(),
  is_internal_tour = col_logical(),
  destination_logsum = col_double(),
  vehicle_occup_1 = col_character(),
  vehicle_occup_2 = col_character(),
  vehicle_occup_3.5 = col_character(),
  tour_mode = col_character(),
  mode_choice_logsum = col_double(),
  selected_vehicle = col_character(),
  atwork_subtour_frequency = col_character(),
  parent_tour_id = col_double(),
  stop_frequency = col_character(),
  primary_purpose = col_character(),
  model = col_character()
))
```

# Reductions 
```{r reductions}
working_df <- tour_df %>%
  select(tour_id, person_id, tour_category, tour_type, stop_frequency) %>%
  group_by(tour_category, tour_type, stop_frequency) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(tour_category, tour_type) %>%
  mutate(share = count/sum(count)) %>%
  ungroup()
```

Examine four selected purposes to see if examples are logical. 

## 1. Work tours
```{r}
summary_df <- working_df %>%
  filter(tour_category == "mandatory" & tour_type == "work") %>%
  arrange(-share)

summary_df %>%
  kbl() %>%
  kable_styling()
```

This makes sense. The dominate outcome is 0 out, 0 in, followed by 0 out, 1 in. It is somewhat illogical that 0 out 3 in is more likely than 0 out, 2 in. The calibration *appears* to have modestly overfit this model to the observed data, but this is a very minor issue. 

## 2. At-work eating tours
```{r}
summary_df <- working_df %>%
  filter(tour_category == "atwork" & tour_type == "eat") %>%
  arrange(-share)

summary_df %>%
  kbl() %>%
  kable_styling()
```

A dominate share of 0 out, 0 in, as expected.

## 3. Shopping tours
```{r}
summary_df <- working_df %>%
  filter(tour_category == "non_mandatory" & tour_type == "shopping") %>%
  arrange(-share)

summary_df %>%
  kbl() %>%
  kable_styling()
```

The 0 out, 0 in alternative is highest, as expected, and the in/out pairs (1 out, 0 in and 0 out, 1 in) are roughly balanced, as expected. 

## 4. Joint social
```{r}
summary_df <- working_df %>%
  filter(tour_category == "joint" & tour_type == "social") %>%
  arrange(-share)

summary_df %>%
  kbl() %>%
  kable_styling()
```

These outcomes appear to be a bit overfit to the data, with 3 out, 0 in being the most likely. Perhaps this is being addressed in the on-going calibration effort? But this is a very minor issue.


# Findings
The models appear to be implemented correctly. Though a minor issue, the models could benefit from calibration refinement (though the impact would be negligible). For example, the base coefficients for stops on social tours on the outbound trip is `1.323` -- it does not vary across the alternatives. During calibration of SANDAG's ABM2, it appears a constant of `3.226` was applied to the `3 out, 0 in` alternative, but a smaller constant of `1.75` to the `2 out, 0 in` alternative. The current calibration adjusts both of these constants down, but not enough to account for the gap. It's unlikely that tours with 3 outbound stops are actually more likely than those with 2. MTC may benefit from cleaning up the UECs prior to calibration. 

ActivitySim files reviewed include:

- `stop_frequency_alternatives.csv`
- `stop_frequency_annotate_tours_preprocessor.csv`
- `stop_frequency_[purpose].csv`, where purpose is `at_work`, `eatout`, `escort`, `othdiscr`, `othmaint`, `school`, `shopping`, `social`, `univ`, and `work`
- `stop_frequency_coefficients_[purpose]`, where purpose is the same as above.
- `stop_frequency.yaml`


###


